The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-12-13 sandbox/derivation.tex
derivation: fix typo
derivation: add powerlaw+constant integrals

\subsubsection{Power law planets (Model \# 3)}

Power law planets are a slightly more realistic mass ratio distribution.
If we now assume
\begin{equation}
    \Gamma_i(r,M) = Z_i \times \frac{r^\delta}{\mathcal{N}_r} \times
    \frac{M^\gamma}{\mathcal{N}_M},
\end{equation}
where $Z_i$ is dimensionless, and the normalization constants carry 
the units (each side has units $[r^{-1} M^{-1}]$).
From Eq.~\ref{eq:simple_apparent_occ_rate}, we can write
\begin{equation}
\Gamma\a(r\a,M\a) = \frac{1}{1+\mu} \left(
\Gamma_0(r\a,M\a) + \frac{{\rm BF}}{1-{\rm BF}} 
\left[ I_1(r_a,M\a) + I_2(r_a,M\a)
\right]
\right),
\end{equation}
where $I_1(r\a,M\a)$ and $I_2(r\a,M\a)$ are defined in the same way as in 
Eqs.~\ref{eq:defn_I1} and~\ref{eq:defn_I2}.
Doing the calculation, we find
\begin{align}
I_1(r\a,M\a) &= 
\int \frac{{\rm d}q}{\mathcal{A}^4} f(q)
\Gamma^1 \left(\frac{r\a}{\mathcal{A}},M\a\right)
\\
&=
\frac{Z_1}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+4}{2}}
\end{align}
and
\begin{align}
I_2(r\a,M\a) &=
\int \frac{q\,{\rm d}q}{\mathcal{A}^3 \mathcal{B}} f(q)
\Gamma^2 \left(\frac{r\a}{\mathcal{B}},qM\a\right)
\\
&= \frac{Z_2}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{8}{3}} 
               (1+q^\alpha)^{\frac{3}{2}}
               (1+q^{-\alpha})^{\frac{\delta+1}{2}}.
\end{align}
The net effect of binarity is to just {\it scale} the overall distribution!

@@

>>>>

\subsubsection{Power law planets (Model \# 6)}

Power law planets are a slightly more realistic mass ratio distribution.
If we now assume
\begin{equation}
    \Gamma_i(r,M) = Z_i \times \frac{r^\delta}{\mathcal{N}_r} \times
    \frac{M^\gamma}{\mathcal{N}_M},
\end{equation}
where $Z_i$ is dimensionless, and the normalization constants carry 
the units (each side has units $[r^{-1} M^{-1}]$).
From Eq.~\ref{eq:simple_apparent_occ_rate}, we can write
\begin{equation}
\Gamma\a(r\a,M\a) = \frac{1}{1+\mu} \left(
\Gamma_0(r\a,M\a) + \frac{{\rm BF}}{1-{\rm BF}} 
\left[ I_1(r_a,M\a) + I_2(r_a,M\a)
\right]
\right),
\end{equation}
where $I_1(r\a,M\a)$ and $I_2(r\a,M\a)$ are defined in the same way as in 
Eqs.~\ref{eq:defn_I1} and~\ref{eq:defn_I2}.
Doing the calculation, we find
\begin{align}
I_1(r\a,M\a) &= 
\int \frac{{\rm d}q}{\mathcal{A}^4} f(q)
\Gamma^1 \left(\frac{r\a}{\mathcal{A}},M\a\right)
\\
&=
\frac{Z_1}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+4}{2}}
\label{eq:I_1_powerlaw_r}
\end{align}
and
\begin{align}
I_2(r\a,M\a) &=
\int \frac{q\,{\rm d}q}{\mathcal{A}^3 \mathcal{B}} f(q)
\Gamma^2 \left(\frac{r\a}{\mathcal{B}},qM\a\right) q^{2/3}
\\
&= \frac{Z_2}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{8}{3}} 
               (1+q^\alpha)^{\frac{3}{2}}
               (1+q^{-\alpha})^{\frac{\delta+1}{2}}.
\label{eq:I_2_powerlaw_r}               
\end{align}
The net effect of binarity is to just {\it scale} the overall distribution!

@@

\subsubsection{Constant rate density of planets}
Combining a power law distribution with a constant below some minimum radius 
would be reasonable (Model \#3).
Here, we'll have
\begin{align}
\notag
\Gamma_i(r,M) &= Z_i \times \frac{r^\delta}{\mathcal{N}_r} \times
\frac{M^\gamma}{\mathcal{N}_M}\quad\quad (r>r_\ell),\\
&= Z_i \times \frac{r_\ell^\delta}{\mathcal{N}_r} \times
\frac{M^\gamma}{\mathcal{N}_M}\equiv c_\ell \quad\quad (r<r_\ell).
\end{align}

Once again,
\begin{equation}
\Gamma\a(r\a,M\a) = \frac{1}{1+\mu} \left(
\Gamma_0(r\a,M\a) + \frac{{\rm BF}}{1-{\rm BF}} 
\left[ I_1(r_a,M\a) + I_2(r_a,M\a)
\right]
\right).
\end{equation}
At $r\a>r_\ell$, the integrals $I_1(r\a,M\a)$ and $I_2(r\a,M\a)$ are given by 
Eqs.~\ref{eq:I_1_powerlaw_r} and~\ref{eq:I_2_powerlaw_r} respectively.
\paragraph{Case $i=1$:}
At $r\a<r_\ell$, they can be evaluated as follows:
\begin{align}
I_1(r\a,M\a) &\equiv \int \frac{{\rm d}q}{\mathcal{A}^4} f(q) 
\Gamma_1\left( \frac{r\a}{\mathcal{A}}, M\a \right) \\
&=
\int_{r\a/\mathcal{A}>r_\ell} \frac{{\rm d}q}{\mathcal{A}^4} f(q) 
\Gamma_1\left( \frac{r\a}{\mathcal{A}}, M\a \right)
+
\int_{r\a/\mathcal{A}<r_\ell} \frac{{\rm d}q}{\mathcal{A}^4} f(q) 
\Gamma_1\left( \frac{r\a}{\mathcal{A}}, M\a \right) \\
&=
\int_{q_\ell(r\a)}^1 \frac{{\rm d}q}{\mathcal{A}^4} f(q) 
\Gamma_1\left( \frac{r\a}{\mathcal{A}}, M\a \right)
+
\int_{0}^{q_\ell(r\a)} \frac{{\rm d}q}{\mathcal{A}^4} f(q) 
\Gamma_1\left( \frac{r\a}{\mathcal{A}}, M\a \right) \\
&=
\frac{Z_1 r\a^\delta M\a^\gamma}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
\int_{q_\ell(r\a)}^1 {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+4}{2}}
+ 
\frac{Z_1 r_\ell^\delta M\a^\gamma}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
\int_0^{q_\ell(r\a)} {\rm d}q\, 
(1+q^\alpha)^2 q^\beta,
\end{align}
where $q_\ell(r\a)$ is found by inverting $r\a/\mathcal{A}=r_\ell$:
\begin{equation}
q_\ell(r\a) = \left(\left(\frac{r_\ell}{r\a}\right)^2 - 
1\right)^{\frac{1}{\alpha}}.
\end{equation}

\paragraph{Case $i=2$:}
For $i=2$, the situation is slightly more complicated. I will specify that 
$\alpha=3.5$, or else the existence of different classes of minima in 
$r\a/\mathcal{B}=r_\ell$ might confuse the matter.
If so, then there are two subcases.
\textit{For subcase \#1,} $\left(\frac{r_\ell}{r\a}\right)^2 > 2$. Then
\begin{align}
I_2(r\a,M\a) &=
\int \frac{q^{5/3}\,{\rm d}q}{\mathcal{A}^3 \mathcal{B}} f(q)
\Gamma^2 \left(\frac{r\a}{\mathcal{B}},qM\a\right) \\
&=
\int_0^{q_{\ell,1}(r\a)} (\ldots)
+
\int_{q_{\ell,1}(r\a)}^{q_t} (\ldots)
+
\int_{q_t}^{1} (\ldots)
\label{eq:intermed_q}\\
&=
\int_0^{q_{\ell,1}(r\a)} (\ldots)
+
\int_{q_{\ell,1}(r\a)}^{1} (\ldots),
\end{align}
where $(\ldots)$ indicates the same integrand as in the first line.
Here, $q_t$ is the turning point of $q^2(1+q^{-\alpha})$ (see 
Figs.~\ref{fig:func_of_q_0} and~\ref{fig:func_of_q_1}),
and $q_{\ell,1}(r\a)$ is the sole root of
\begin{equation}
q^2 (1+q^{-\alpha}) = \left(\frac{r_\ell}{r\a}\right)^2,
\end{equation} 
which must be found numerically.
In Eq~\ref{eq:intermed_q}, the left-most term has $r\a/\mathcal{B} > r_\ell$, 
while the right two terms have $r\a/\mathcal{B} < r_\ell$ (and so they can be 
safely merged).
Simplifying,
\begin{align}
\notag
I_2(r\a,M\a) &= 
    \frac{1}{\mathcal{N}_q}
    \int_0^{q_{\ell,1}(r\a)} 
    {\rm d}q\, q^{8/3} (1+q^\alpha)^{3/2} q^\beta
    \Gamma^2 \left(\frac{r\a}{\mathcal{B}},qM\a\right)
    \\
&\quad\quad\quad+
\frac{1}{\mathcal{N}_q}
\int_{q_{\ell,1}(r\a)}^{1}
    {\rm d}q\, q^{8/3} (1+q^\alpha)^{3/2} q^\beta
    \Gamma^2 \left(\frac{r\a}{\mathcal{B}},qM\a\right),
\end{align}
and plugging in the rate densities, we get
\begin{align}
\notag
I_2(r\a,M\a) &= 
\frac{Z_2 r\a^\delta M\a^\gamma}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M}
\int_0^{q_{\ell,1}(r\a)} 
{\rm d}q\,
q^{\beta+\gamma+\delta+\frac{8}{3}} 
(1+q^\alpha)^{\frac{3}{2}}
(1+q^{-\alpha})^{\frac{\delta+1}{2}}
\\
&\quad\quad\quad+
\frac{Z_2 r_\ell^\delta M\a^\gamma}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M}
\int_{q_{\ell,1}(r\a)}^{1}
{\rm d}q\, q^{\beta + \gamma + \frac{8}{3}} (1+q^\alpha)^{3/2} 
(1+q^{-\alpha})^{\frac{1}{2}}.
\quad\quad \left({\rm for}\ \left(\frac{r_\ell}{r\a}\right)^2 > 2\right)
\end{align}

\textit{For subcase \#2,} if $q_t^2 (1+q_t^{-\alpha}) < 
\left(\frac{r_\ell}{r\a}\right)^2 < 2 $,
then there are two roots, $q_{\ell,1}$ (which is $<q_t$) and $q_{\ell,2}$ 
(which is $>q_t$). Each must be found 
numerically. Their existence makes the integrals silly:
\begin{align}
I_2(r\a,M\a) &=
\int \frac{q^{5/3}\,{\rm d}q}{\mathcal{A}^3 \mathcal{B}} f(q)
\Gamma^2 \left(\frac{r\a}{\mathcal{B}},qM\a\right) \\
&=
\int_0^{q_{\ell,1}(r\a)} (\ldots)
+
\int_{q_{\ell,1}(r\a)}^{q_t} (\ldots)
+
\int_{q_t}^{q_{\ell,2}(r\a)} (\ldots)
+
\int_{q_{\ell,2}(r\a)}^1 (\ldots)
\\
&=
\int_0^{q_{\ell,1}(r\a)} (\ldots)
+
\int_{q_{\ell,1}(r\a)}^{q_{\ell,2}(r\a)} (\ldots)
+
\int_{q_{\ell,2}(r\a)}^1 (\ldots),
\end{align}
where the left and right-most terms have $r\a/\mathcal{B}>r_\ell$, and the 
center term has $r\a/\mathcal{B}<r_\ell$.
Numerically, $q_t = 0.9210924370$ and 
$q_t^2(1+q^{-\alpha})=1.9796263301$.

Plugging in the rate densities, we get
\begin{align}
\notag
I_2(r\a,M\a) &=
\frac{Z_2 r\a^\delta M\a^\gamma}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M}
\int_0^{q_{\ell,1}(r\a)} 
{\rm d}q\,
q^{\beta+\gamma+\delta+\frac{8}{3}} 
(1+q^\alpha)^{\frac{3}{2}}
(1+q^{-\alpha})^{\frac{\delta+1}{2}}
\\
\notag
&\quad\quad\quad+
\frac{Z_2 r_\ell^\delta M\a^\gamma}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M}
\int_{q_{\ell,1}(r\a)}^{q_{\ell,2}(r\a)}
{\rm d}q\, q^{\beta + \gamma + \frac{8}{3}} (1+q^\alpha)^{3/2} 
(1+q^{-\alpha})^{\frac{1}{2}}
\\
\notag
&\quad\quad\quad+
\frac{Z_2 r\a^\delta M\a^\gamma}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M}
\int_{q_{\ell,2}(r\a)}^1
{\rm d}q\,
q^{\beta+\gamma+\delta+\frac{8}{3}} 
(1+q^\alpha)^{\frac{3}{2}}
(1+q^{-\alpha})^{\frac{\delta+1}{2}}
\\
&\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
\ldots\ {\rm for}\ q_t^2 
(1+q_t^{-\alpha}) < 
\left(\frac{r_\ell}{r\a}\right)^2 < 2.
\end{align}

This should actually be enough to integrate semi-analytically with a computer.

\begin{figure}[!th]
    \centering
    \includegraphics[width=.7\textwidth]{funcq_of_q_0.png}
    \caption{For $\alpha=3.5$, there is a single minimum of $q^2 
    (1+q^{-\alpha})$.}
    \label{fig:func_of_q_0}
\end{figure}
\begin{figure}[!th]
    \centering
    \includegraphics[width=.7\textwidth]{funcq_of_q_1.png}
    \caption{For $\alpha=3.5$, there is a single minimum of $q^2 
        (1+q^{-\alpha})$.}
    \label{fig:func_of_q_1}
\end{figure}
